\chapter{Conclusions}
\label{chap:conclusions}

\section{Summary of Work Undertaken}

This project set out to design, implement, and evaluate a scalable Agentic AI Tutor built from subject-specific Small Language Models, deployed using MLOps practices. The core development lifecycle was completed across six work packages, from data preparation through deployment, CI/CD automation, and multimodal output integration.

Three LoRA adapters were fine-tuned on Phi-4-mini-instruct using subject-specific question-answer datasets derived from institutional lecture materials for Fundamentals of Software Development (FSD), Fundamentals of Computer Science (FCS), and Discrete Mathematics (DMA). Each adapter was paired with a dedicated ChromaDB vector store populated with course slides and tutorials, forming a Retrieval-Augmented Generation pipeline that grounds model responses in verifiable course content.

The complete system was containerised as a microservices architecture using Docker Compose and Kubernetes, with three independent backend services, three ChromaDB instances, and a Next.js frontend. A GitOps delivery pipeline was established using GitHub Actions and Argo CD, providing automated linting, security scanning, end-to-end testing, and image publishing on every push to main branch.

Accessibility features were implemented through a multimodal output layer, including browser-based text-to-speech synthesis and a real-time talking-head avatar. The system uses HeadTTS for speech generation and TalkingHead for avatar animation, enabling low-latency, voice-based interaction for diverse learners without relying on backend media generation services.

\section{Critical Evaluation of Requirements}

The five objectives stated in the Introduction are evaluated in turn.

\textbf{Objective 1: Apply MLOps practices.} Version control, automated CI checks, reproducible builds, container publishing, and GitOps-based delivery were implemented. Python syntax checking, Bandit security scanning, npm audit, Playwright smoke tests, and Trivy image scanning are integrated into GitHub Actions, while Argo CD provides automated sync and self-healing. Model adapter versions are tracked in subjects.yaml and exposed through /health. This objective is fully met.

\textbf{Objective 2: Adapt one base SLM into three subject-specific tutor variants.} All three LoRA adapters were successfully trained and evaluated on top of Phi-4-mini-instruct. Fine-tuning consistently improved BLEU-2, ROUGE-L, and BERTScore compared to the base model across all subjects, confirming that subject-specific adaptation is effective. This objective is fully met.

\textbf{Objective 3: Integrate a RAG pipeline.} RAG was implemented and evaluated, producing further gains in BLEU-2 across all three subjects and marginal ROUGE-L improvements in FSD and DMA. The drop in METEOR is interpreted as a consequence of domain-specific phrasing rather than degraded quality, a conclusion supported by simultaneous BERTScore improvement. The pipeline also exposes retrieved source chunks to users, directly fulfilling the explainability requirement. This objective is fully met.

\textbf{Objective 4: Fine-tune, evaluate, and benchmark SLM variants with quality and runtime metrics.} Quantitative evaluation was conducted using BLEU-2, ROUGE-L, METEOR, BERTScore F1, and runtime indicators (mean latency, tokens-per-second, and mean generated output length) across all three subjects, with comparison against larger LLM baselines. This objective is fully met.

\textbf{Objective 5: Deploy the system as containerised, scalable services with an optional multimodal interface.} The system was deployed as a containerised microservices architecture using Docker Compose and Kubernetes, with three subject-specific backend services, three ChromaDB instances, and a Next.js frontend. The architecture provides isolation and fault containment per subject and is designed for scalable deployment. An optional multimodal interface was implemented in the browser using HeadTTS for speech synthesis and TalkingHead for real-time avatar animation, enabling low-latency audio-avatar interaction without backend media generation services. A practical limitation in the local WSL2 Kubernetes setup was the absence of GPU passthrough, so local cluster inference ran on CPU. This is a local platform constraint rather than an architectural deficiency. This objective is substantially met.


\section{Critical Evaluation of Project Goals}

The central aim was to demonstrate that subject-specific SLMs fine-tuned with LoRA and augmented with RAG can deliver accurate, low-latency, and cost-effective tutoring responses as an alternative to large commercial models.

Latency values reported in this dissertation were obtained in a dedicated benchmarking environment and should be interpreted as setup-specific runtime measurements.

The evaluation results support this claim. On BLEU-2 and BERTScore F1, the adapted SLMs outperform both LLM baselines (Llama-3.3-70B and Qwen3-32B) across all three subjects, demonstrating that domain adaptation closes the performance gap that would otherwise exist between a 3.8-billion-parameter model and models ten to twenty times larger. Inference latency for the SLMs (mean 4-5 seconds) is approximately half that of the Qwen3-32B baseline, supporting the claim that SLMs are suitable for low-latency online learning environments. The modular architecture allows each subject to be updated, retrained, or scaled without affecting the others, satisfying the requirement for independent subject modules.

The main quantitative limitation is METEOR: the LLMs, particularly Llama-3.3-70B, score higher on this metric across all subjects. As discussed in the Analysis section, this is largely an artefact of how METEOR penalises fragmented word-level alignment. When a fine-tuned SLM produces a correct but differently phrased answer, METEOR may score it lower than a more generic base model that happens to partially echo the reference phrasing. The BERTScore results, which measure semantic rather than lexical similarity, do not exhibit this reversal, which is the more meaningful indicator for tutoring quality.

Overall, the project demonstrates that a specialised SLM system can match or exceed the output quality of much larger general-purpose LLMs in a subject-specific domain, while operating at lower latency and without reliance on commercial API services.

\section{Reflection on the Original Proposal}

Several decisions taken during implementation diverged from the original plan in ways that improved the final system.

The original monitoring plan (WP5) proposed Prometheus and Grafana dashboards. These were not implemented because the effort required to configure and integrate observability infrastructure within the remaining project timeline would have reduced time available for evaluation and writing. Structured logging to stdout, health check endpoints, and request-level inference logging were implemented as a practical alternative, providing sufficient observability for a single-developer deployment. Prometheus integration remains the most significant item of remaining work.


\section{Future Work}

Several directions are identified for future development.

\textbf{Real-time monitoring.} The most important unimplemented component is a Prometheus metrics endpoint and Grafana dashboard. Adding these would enable detection of model drift, retrieval quality degradation, and latency anomalies in production, completing the WP5 objectives.

\textbf{Continuous retraining pipeline.} Currently, retraining a LoRA adapter requires manual execution of the Colab notebook. Automating this as a triggered Argo Workflow when course materials are updated would complete the lifecycle management vision of WP6 and allow the system to remain current without manual intervention.

\textbf{User feedback and drift detection.} Logging anonymised question-answer pairs and student satisfaction signals would enable evaluation of model quality over time and inform targeted retraining. A structured feedback loop would also support the teacher-in-the-loop validation approach described in the ethical considerations.

\textbf{Cloud GPU deployment.} The Kubernetes deployment model can be extended for cloud GPU environments. In practice, this requires adding explicit GPU resource requests/limits and node scheduling constraints (e.g., for EKS GPU node groups). Deploying to a managed service such as AWS EKS with GPU node pools would validate the production scalability of the architecture and remove the WSL2 GPU passthrough limitation encountered locally.

\textbf{Load and stress testing with JMeter.} A dedicated performance-testing stage can be added to validate behaviour under concurrent user load. This should include p95/p99 latency, throughput, and error-rate targets, and be integrated into the release process for production readiness.

\textbf{Extending to additional subjects.} The modular architecture was explicitly designed to support new subjects with minimal changes. Adding a new subject requires preparing a QA dataset, training a LoRA adapter, populating a ChromaDB vector store, and adding one service entry to the Kubernetes manifests. This extensibility could be evaluated in a follow-on project.

\textbf{University authentication and registration.} The system currently has no authentication layer. Integrating institutional single sign-on (OAuth 2.0 or SAML) and account registration/profile management would restrict access to authorised students, satisfy access-control requirements, and enable user-level personalisation. Authentication would also provide the user identity needed to associate persistent conversation history with individual accounts.

\textbf{Conversation memory and persistent history.} The current implementation already supports short-term multi-turn context by passing recent message history in each request. The main remaining gap is persistent, account-linked history across sessions and devices. Storing conversation history in a backend database and associating it with authenticated users would allow session continuity, longitudinal learning support, and better insight into recurring student difficulties.



\section{Personal Reflection}

Before this project, I had no hands-on experience with training machine learning models. 
Working on this dissertation gave me my first end-to-end exposure to the full workflow: data preparation, adapter training, evaluation, deployment, and iterative improvement.

I learned how retrieval quality, prompt design, API behaviour, and frontend interaction all influence the final user experience. 
This helped me move from a purely theoretical understanding of AI to a practical engineering perspective.

If I repeated the work, I would implement authentication earlier in the timeline, so that account-linked conversation history could be evaluated as part of the core system rather than deferred to future work.

Overall, this project marked a major step in my development: from having no prior model-training experience to being able to build and evaluate a complete AI tutor pipeline with clear technical and methodological justification.